___
title: "Exploring the Impact of Transmission Type on Automobile Efficiency"
author: "Logan J Travis"
date: "27 July 2014"
output: html_document
---

```{r setup, echo = FALSE}
# Load libraries
library(datasets)

# Define global variables
dat <- mtcars
dat$am.factor <- factor(dat$am, labels = c("Automatic", "Manual"))
```

# Exploring the Impact of Transmission Type on Automobile Efficiency
## Logan J Travis
## 27 July 2014

### Executive Summary

### Exploratory Analysis
* Briefly describe mtcars data citing source in appendix
* Outline methodology including ploting techniques and ...

#### Single Linear Regression Using Transmission Type
[box plot MPG by transmisison type](#exPlot01)

[scatter plot with linear model fit MPG ~ transmission type](#exPlot02)

[scatter plot with linear model fit MPG ~ weight (1,000lbs)](#exPlot03)

#### Comparison to Competing Predictors
Table (maybe in appendix)

### Multi-Variable Regression (CHANGE TITLE)
* Briefly recap exploratory analysis
* Plot data, fit, confidence, prediction, and residuals
* Create an 'inference'

### Conclusion

### Appendices
#### A: Figures
<ol>
    <li><a name = "exPlot01"></a>
        ``` {r exPlot01, echo = FALSE,}
        with(dat, plot(am.factor, mpg, xLab = "Transmittion", yLab = "MPG"))
        title("MPG Comparison by Transmission Type")
        ```
    </li>
    <li><a name = "exPlot02"></a>
        ``` {r exPlot02, echo = FALSE}
        # Fit linear model mpg ~ transmission
        fit <- lm(mpg ~ am, dat)
        fitCoef <- summary(fit)$coefficients
        fitText <- c(paste("Intercept:", round(fitCoef[1, 1], digits = 4)),
                     paste("Intercept P-Value:", signif(fitCoef[1, 4], digits = 4)),
                     paste("Slope:", round(fitCoef[2, 1], digits = 4)),
                     paste("Slope P-Value:", signif(fitCoef[2, 4], digits = 4)))
        
        # Scatter plot mpg ~ transmission
        # Fits linear model line
        with(dat, plot(am, mpg, xlab = "Transmission Type", ylab = "MPG", xaxt = "n"))
        abline(fit, col = "red", lwd = 2)
        axis(1, at = 0:1, labels = c("Automatic", "Manual"))
        text(x = 0.65, y = seq(from = 32, to = 26, by = -2), adj = c(1, 1), labels = fitText)
        title("MPG ~ Transmission Type Linear Model Fit")
        
        # Remove local variables
        rm(fit, fitCoef, fitText)
        ```
    </li>
    <li><a name = "exPlot03"></a>
        ```{r exPlot03, echo = FALSE, }
        # Fit linear model mpg ~ transmission
        fit <- lm(mpg ~ wt, dat)
        fitCoef <- summary(fit)$coefficients
        fitText <- c(paste("Intercept:", round(fitCoef[1, 1], digits = 4)),
                     paste("Intercept P-Value:", signif(fitCoef[1, 4], digits = 4)),
                     paste("Slope:", round(fitCoef[2, 1], digits = 4)),
                     paste("Slope P-Value:", signif(fitCoef[2, 4], digits = 4)))
        
        # Scatter plot mpg ~ weight
        # Points colord for blue = "Automatic"" and red = "Manual"
        # Fits linear model
        with(dat, plot(wt, mpg, xlab = "Weight (1,000lbs)", ylab = "MPG", type = "n"))
        with(subset(dat, am == 0), points(wt, mpg, col = "blue"))
        with(subset(dat, am == 1), points(wt, mpg, col = "red"))
        abline(fit, lwd = 2)
        legend("topright", pch = 1, col = c("blue", "red"), legend = c("Automatic", "Manual"))
        text(x = 4.7, y = seq(from = 34, to = 27, by = -2), adj = c(1, 1), labels = fitText)
        title("MPG ~ Weight (1,000lbs) Linear Model Fit")
        
        # Remove local variables
        rm(fit, fitCoef, fitText)
        ```
    </li>
</ol>

#### B: Shiny App