---
title       : Exploring Transmission Impact on Automobile Efficiency
subtitle    : Using Shiny App and MTCARS Dataset
author      : Logan J Travis
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Executive Summary
* Many struggle to understand the impact of discrete variables in statistical models
* Investigate automobile efficiency as a familiar and accessible example
* Design a simple application targeting non-statisticians to:
    * Interactively model fuel efficiency (MPG) using a variety of continuous variables<br>(e.g. weight, engine displacement)
    * Compare model fit after adding a discrete variable (transmission type)
* Host the application online for all to experience!

---

## Dataset and Methodology
### [MTCARS Dataset](http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html)
* Part of the base R package
* Extracted from the 1974 *Motor Trend* US magazine
* Compares fuel consumption against 10 variable for 32 automobiles

### Linear Regression Models
* Regression by transmission type yields useful results (details on next slide)
* Attempt to improve by regressing other variables
* Compare single variable versus adding transmission type

--- &twocol

## Example Regression Against Transmission Type

*** =left

```{r plotMPG~Trans, echo = FALSE}
    # Load libraries
    library(datasets)
    
    # Read 'mtcars' dataset
    dat <- mtcars
    
    # Fit linear model for mpg ~ am
    fit <- lm(mpg ~ am, dat)
    fitCoef <- summary(fit)$coefficients
    fitText <- c(paste("Intercept:", round(fitCoef[1, 1], digits = 4)),
                 paste("Intercept P-Value:", signif(fitCoef[1, 4], digits = 4)),
                 paste("Slope:", round(fitCoef[2, 1], digits = 4)),
                 paste("Slope P-Value:", signif(fitCoef[2, 4], digits = 4)))
    
    # Scatter plot mpg ~ am
    # Adds line for 'fit'
    plot(mpg ~ am, dat, xlab = "", ylab = "MPG", xaxt = "n")
    abline(fit, lwd = 2)

    # Add labels for summary data, axes, and title
    text(x = 0.65, y = seq(from = 32, to = 26, by = -2), adj = c(1, 1), labels = fitText)
    axis(1, c(0, 1), c("Automatic", "Manual"))
    title("MTCARS Regression MPG ~ Transmission Type")
```

*** =right

* Linear regression for MPG ~ transmission *suggests* a benefit for manuals
    * The P-value for the slope (two-sided) bests a 99.9% confidence interval
    * The P-value for the intercept greatly exceeds 99.9% confidence
* Yet, predictions from transmission alone spread +/- 10 MPG
* Would you buy a car offering 6 to 27 MPG?

```{r summaryMPG~Trans, echo = FALSE}
    # Load libraries
    library(datasets)
    
    # Read 'mtcars' dataset
    dat <- mtcars

    # Fit linear model for mpg ~ am
    fit <- lm(mpg ~ am, dat)

    # Create data.frame for unique transmission numbers
    trans <- data.frame(am = unique(dat$am))

    # Bind columns for prediction interval at 95%
    trans <- cbind(trans, predict(fit, trans, interval = "prediction"))
    
    # Transform transmission number to type and re-label
    trans <- transform(trans, am = factor(trans$am, labels = c("Automatic", "Manual")))
    names(trans) <- c("Transmission", "Fit", "Lower", "Upper")

    # Return data.frame
    trans
```

---

## Shiny App

* Available on the [Rstudio Shiny App Server](http://lonejt.shinyapps.io/shiny_app/)
* Input controls for predictor, interval, and color coding by transmission

    <img src = "https://raw.githubusercontent.com/l0neJT/reg_models01/master/slidify_presentation/assets/img/app_inputs.png" />

* Plots comparing single and multi-variable regression
* Tabs for summary statistics and table data

---

## Interpreting Results

* Many variables yield tighter fits than transmission type notably:
    * Displacement (cu. in.) with P-value of 9.38e-10
    * Weight (lb/1000) with a P-value of 1.29e-10
* Adding transmission type did not improve better-fit variables
* Further modeling - perhaps adjusting for covariance - may yield improved predictions